SI618 Group Project
========================================================

Stat Package Format Analysis
--------------------------------------------------------

### All statistical package formats used overall, color-coded by country:

```{r echo=FALSE}
library(DBI)
library(RSQLite)
library(ggplot2)
library(plyr)


# Connect to the database...
driver<-dbDriver("SQLite")
connect<-dbConnect(driver, dbname = "../db/weblogs.db")

# Retrieve all weblog entries for stat package reports and define factors...
log.entries = dbGetQuery(connect, "
select wl.timestamp, u.name as institution, c.name as country, c.abbr as abbr, wl.study, wl.browser_type,
    case
        when wl.format like '%spss%' then 'SPSS'
        when wl.format like '%stata%' then 'STATA'
        when wl.format like '%sas%' then 'SAS'
        when wl.format = 'ascii' or wl.format = 'delimited' or wl.format = 'other' then 'DELIMITED'
        else wl.format
    end as format,
    case
        when (browser_type like 'Mozilla%' or browser_type like 'Opera%' or browser_type like 'Safari%' or browser_type like 'Nokia%' or browser_type like 'BlackBerry%') and
             (browser_type not like '%Yandex%' and browser_type not like '%spider%' and browser_type not like '%Babya Discover%'
              and browser_type not like '%bot%' and browser_type not like '%Daumoa%' and browser_type not like '%crawler%' and browser_type not like '%LinkCheck%'
              and browser_type not like '%WebCorp%' and browser_type not like '%Yahoo! Slurp%' and browser_type not like '%Genieo%') then 'Human'
        when browser_type like '%CopperEgg%' or browser_type like '%InterMapper%' then 'ICPSR-Bot'
        else 'Bot'
    end as accessor
from weblog_downloads wl left outer join country_ips ci on wl.long_ip = ci.long_ip
                         left outer join countries c on ci.country_id = c.id
                         left outer join university_ips ui on (wl.abyte = ui.abyte and wl.bbyte = ui.bbyte and wl.cbyte = ui.cbyte and wl.dbyte >= ui.dbyte1 and wl.dbyte <= ui.dbyte2)
                         left outer join universities u on ui.university_id = u.id
")

log.entries$country <- factor(log.entries$country)
log.entries$abbr <- factor(log.entries$abbr)
log.entries$format <- factor(log.entries$format)
log.entries$accessor <- factor(log.entries$accessor)
log.entries$institution <- factor(log.entries$institution)

# All statistical packages used overall, color-coded by country...
ggplot(log.entries, aes(format, colour = abbr, fill = abbr)) +
        geom_histogram() + ylab("Count") + xlab("Format") +
        labs(title = "Overall Downloads by Format") + 
        guides(col=guide_legend(ncol=3, byrow=T))
```

### All statistical packages by country with NA removed:

```{r echo=FALSE}
# All statistical packages by country with NA removed...
log.entries.nona = subset(log.entries, !(is.na(abbr)))
ggplot(log.entries.nona, aes(format, colour = abbr, fill = abbr)) +
  geom_histogram() + ylab("Count") + xlab("Format") +
  labs(title = "Download Format by Country") + 
  guides(col=guide_legend(ncol=3, byrow=T))
```

### All statistical package formats by country with NA and 'US' removed:

```{r echo=FALSE}

# All statistical packages by country with NA and 'US' removed...
log.entries.nous = subset(log.entries.nona, abbr != 'US')
ggplot(log.entries.nous, aes(format, colour = abbr, fill = abbr)) +
  geom_histogram() + ylab("Count") + xlab("Format") +
  labs(title = "Download Format by Country ('US' Excluded)") + 
  guides(col=guide_legend(ncol=3, byrow=T))
```
### Top 5 statistical package format downloads by country

```{r echo=FALSE}
# Listing top 5 statistical package downloads by country...
stat.package.summary = ddply(log.entries.nona, c("country", "format"), summarise,
                             count = length(country))
stat.package.summary = stat.package.summary[order(stat.package.summary$count,
                                                  decreasing= TRUE),]
for (f in levels(log.entries$format)) {
  stat.package.summary.format = subset(stat.package.summary, format == f)
  stat.package.summary.format = head(stat.package.summary.format, 5)
  stat.package.summary.format <- stat.package.summary.format[
    order(stat.package.summary.format$count, decreasing = TRUE),]
  stat.package.summary.format$country <- reorder(
    stat.package.summary.format$country, -stat.package.summary.format$count)
  plot.title <- paste("Top Downloaders of ", f, sep="")
  print(ggplot(stat.package.summary.format, aes(country, count)) +
          geom_histogram(stat="identity") + ylab("Count") + xlab("Country") +
          labs(title = plot.title) +
          theme(axis.text.x=element_text(angle=45, hjust=1)))
  idx <- 1
  for (c in stat.package.summary.format$country) {
    log.entries.subset = subset(log.entries, country == c)
    plot.title <- paste("Top Downloaders of ", f, " (", idx, "):  ", c, sep="")
    print(ggplot(log.entries.subset, aes(format, colour = format, fill = format)) +
            geom_histogram() + ylab("Count") + xlab("Format") +
            labs(title = plot.title))
    idx <- idx+1
  }
}
```

### Top 5 statistical package format downloads by institution:

```{r echo=FALSE}
# Listing top 5 statistical package downloads by institution...
log.entries.nona.inst = subset(log.entries, !(is.na(institution)))
stat.package.summary = ddply(log.entries.nona.inst, c("institution", "format"), summarise, count = length(institution))
stat.package.summary = stat.package.summary[order(stat.package.summary$count, decreasing= TRUE),]
for (f in levels(log.entries$format)) {
  stat.package.summary.format = subset(stat.package.summary, format == f)
  stat.package.summary.format = head(stat.package.summary.format, 5)
  stat.package.summary.format <- stat.package.summary.format[
    order(stat.package.summary.format$count, decreasing = TRUE),]
  stat.package.summary.format$institution <- reorder(
    stat.package.summary.format$institution, -stat.package.summary.format$count)
  plot.title <- paste("Top Downloaders of ", f, sep="")
  print(ggplot(stat.package.summary.format, aes(institution, count)) +
          geom_histogram(stat="identity") + ylab("Count") + xlab("Institution") +
          labs(title = plot.title) +
          theme(axis.text.x=element_text(angle=45, hjust=1)))

  idx <- 1
  for (i in stat.package.summary.format$institution) {
    log.entries.subset = subset(log.entries, institution == i)
    plot.title <- paste("Top Downloaders of ", f, " (", idx, "):  ", i, sep="")
    print(ggplot(log.entries.subset, aes(format, colour = format, fill = format)) +
            geom_histogram() + ylab("Count") + xlab("Format") +
            labs(title = plot.title))
    idx <- idx+1
  }
}
```

### Bots v humans:

```{r echo=FALSE}
# Breakdown into bots v humans...
for (a in levels(log.entries$accessor)) {
  log.entries.accessor = subset(log.entries.nona, accessor == a)
  plot.title <- paste("Formats Preferred by ", a, "s", sep="")
  print(ggplot(log.entries.accessor, aes(format, colour = format, fill = format)) +
          geom_histogram() + ylab("Count") + xlab("Format") +
          labs(title = plot.title))

  stat.package.summary = ddply(log.entries.accessor, c("country", "format"),
                               summarise, count = length(country))
  stat.package.summary = stat.package.summary[order(stat.package.summary$count,
                                                    decreasing= TRUE),]
  for (f in levels(log.entries.accessor$format)) {
    stat.package.summary.format = subset(stat.package.summary, format == f)
    stat.package.summary.format = head(stat.package.summary.format, 5)
    stat.package.summary.format <- stat.package.summary.format[
      order(stat.package.summary.format$count, decreasing = TRUE),]
    stat.package.summary.format$country <- reorder(
      stat.package.summary.format$country, -stat.package.summary.format$count)
    plot.title <- paste("Top Downloaders of ", f, " by ", a, "s", sep="")
    print(ggplot(stat.package.summary.format, aes(country, count)) +
            geom_histogram(stat="identity") + ylab("Count") + xlab("Country") +
            labs(title = plot.title) +
            theme(axis.text.x=element_text(angle=45, hjust=1)))
    
    idx <- 1
    for (c in stat.package.summary.format$country) {
      log.entries.subset = subset(log.entries.accessor, country == c)
      plot.title <- paste("Top Downloaders of ", f, " (", idx, ") by ", a, "s in ",
                          c, sep="")
      print(ggplot(log.entries.subset, aes(format, colour = format, fill = format))
            + geom_histogram() + ylab("Count") + xlab("Format") +
              labs(title = plot.title))
      idx <- idx+1
    }
  }
}
```
