ICPSR Weblog Analysis - Bots v Humans
========================================================

Chi-square Independence Test Analysis
--------------------------------------------------------

### Statistical Package Format and Accessor (Bots/Humans)

This chi-square test verifies whether or not there is a statistical association between the statistical package format and a bot/human accessor for download data.

H<sub>0</sub>:  There **is no** association between statistical package format and accessor in the population of downloads in ICPSR web logs (they are independent).

H<sub>A</sub>:  There **is** an association between statistical package format and accessor in the population of downloads in ICPSR web logs.

```{r echo=FALSE}
library(DBI)
library(RSQLite)
library(stringr)
library(plyr)
library(ggplot2)

# Connect to the database...
driver<-dbDriver("SQLite")
connect<-dbConnect(driver, dbname = "../db/weblogs.db")

# Retrieve all weblog entries for stat package reports and define factors...
log.entries = dbGetQuery(connect, "
select wl.timestamp, wl.study, wl.browser_type, c.description || ',' as classification,
  case
    when wl.format like '%spss%' then 'SPSS'
    when wl.format like '%stata%' then 'STATA'
    when wl.format like '%sas%' then 'SAS'
    when wl.format = 'ascii' or wl.format = 'delimited' or wl.format = 'other' then 'DELIMITED'
    else wl.format
  end as format,
  case
    when (browser_type like 'Mozilla%' or browser_type like 'Opera%' or browser_type like 'Safari%' or browser_type like 'Nokia%' or browser_type like 'BlackBerry%') and
         (browser_type not like '%Yandex%' and browser_type not like '%spider%' and browser_type not like '%Babya Discover%'
         and browser_type not like '%bot%' and browser_type not like '%Daumoa%' and browser_type not like '%crawler%' and browser_type not like '%LinkCheck%'
         and browser_type not like '%WebCorp%' and browser_type not like '%Yahoo! Slurp%' and browser_type not like '%Genieo%') then 'Human'
         when browser_type like '%CopperEgg%' or browser_type like '%InterMapper%' then 'ICPSR-Bot'
    else 'Bot'
  end as accessor
from weblog_downloads wl inner join classifications c on wl.study = c.study_id
")

log.entries$format <- factor(log.entries$format)
log.entries$accessor <- factor(log.entries$accessor)
log.entries$classification <- factor(log.entries$classification)

# Perform a chi-square test on format vs accessor, report results...
ct <- chisq.test(log.entries$format, log.entries$accessor)
ct
```

Reject H<sub>0</sub>.  There is statistically significant evidence to support an association between statistical package format and accessor in the population of downloads in ICPSR web logs.  Residual value summary:

```{r echo=FALSE}
# Graph residuals...
res <- as.data.frame(ct$residuals)
the.plot <- ggplot(res, aes(log.entries.format, Freq, color = log.entries.accessor, group = log.entries.accessor)) +
  geom_line() + ylab("Residual") + xlab("Download Format") +
  guides(col=guide_legend(title = "Accessor"))
the.plot + labs(title = "Residuals of Format v Accessor")
the.plot + labs(title = "Residuals of Format v Accessor (Scale-adjusted)") + scale_y_continuous(limits = c(-100, 100))
```

Stat Package Format Histograms
--------------------------------------------------------

### Bots v Humans Histograms:

First, creates a summary histogram showing the formats most downloaded by accessor (bots and humans).  Then, further breaks the data down by creating a histogram showing the top 5 countries who downloaded each format for each accessor, along with a histogram showing the breakdown of download formats for each of these top 5 country accessors.

```{r echo=FALSE}

# Retrieve all weblog entries for stat package reports and define factors...
log.entries = dbGetQuery(connect, "
select wl.timestamp, u.name as institution, c.name as country, c.abbr as abbr, wl.study, wl.browser_type,
    case
        when wl.format like '%spss%' then 'SPSS'
        when wl.format like '%stata%' then 'STATA'
        when wl.format like '%sas%' then 'SAS'
        when wl.format = 'ascii' or wl.format = 'delimited' or wl.format = 'other' then 'DELIMITED'
        else wl.format
    end as format,
    case
        when (browser_type like 'Mozilla%' or browser_type like 'Opera%' or browser_type like 'Safari%' or browser_type like 'Nokia%' or browser_type like 'BlackBerry%') and
             (browser_type not like '%Yandex%' and browser_type not like '%spider%' and browser_type not like '%Babya Discover%'
              and browser_type not like '%bot%' and browser_type not like '%Daumoa%' and browser_type not like '%crawler%' and browser_type not like '%LinkCheck%'
              and browser_type not like '%WebCorp%' and browser_type not like '%Yahoo! Slurp%' and browser_type not like '%Genieo%') then 'Human'
        when browser_type like '%CopperEgg%' or browser_type like '%InterMapper%' then 'ICPSR-Bot'
        else 'Bot'
    end as accessor
from weblog_downloads wl left outer join country_ips ci on wl.long_ip = ci.long_ip
                         left outer join countries c on ci.country_id = c.id
                         left outer join university_ips ui on (wl.abyte = ui.abyte and wl.bbyte = ui.bbyte and wl.cbyte = ui.cbyte and wl.dbyte >= ui.dbyte1 and wl.dbyte <= ui.dbyte2)
                         left outer join universities u on ui.university_id = u.id
")

log.entries$country <- factor(log.entries$country)
log.entries$abbr <- factor(log.entries$abbr)
log.entries$format <- factor(log.entries$format)
log.entries$accessor <- factor(log.entries$accessor)
log.entries$institution <- factor(log.entries$institution)
log.entries.nona = subset(log.entries, !(is.na(abbr)))

# Breakdown into bots v humans...
for (a in levels(log.entries$accessor)) {
  log.entries.accessor = subset(log.entries.nona, accessor == a)
  plot.title <- paste("Formats Preferred by ", a, "s", sep="")
  print(ggplot(log.entries.accessor, aes(format, colour = format, fill = format)) +
          geom_histogram() + ylab("Count") + xlab("Format") +
          labs(title = plot.title))

  stat.package.summary = ddply(log.entries.accessor, c("country", "format"),
                               summarise, count = length(country))
  stat.package.summary = stat.package.summary[order(stat.package.summary$count,
                                                    decreasing= TRUE),]
  for (f in levels(log.entries.accessor$format)) {
    stat.package.summary.format = subset(stat.package.summary, format == f)
    stat.package.summary.format = head(stat.package.summary.format, 5)
    stat.package.summary.format <- stat.package.summary.format[
      order(stat.package.summary.format$count, decreasing = TRUE),]
    stat.package.summary.format$country <- reorder(
      stat.package.summary.format$country, -stat.package.summary.format$count)
    plot.title <- paste("Top Downloaders of ", f, " by ", a, "s", sep="")
    print(ggplot(stat.package.summary.format, aes(country, count)) +
            geom_histogram(stat="identity") + ylab("Count") + xlab("Country") +
            labs(title = plot.title) +
            theme(axis.text.x=element_text(angle=45, hjust=1)))
    
    idx <- 1
    for (c in stat.package.summary.format$country) {
      log.entries.subset = subset(log.entries.accessor, country == c)
      plot.title <- paste("Top Downloaders of ", f, " (", idx, ") by ", a, "s in ",
                          c, sep="")
      print(ggplot(log.entries.subset, aes(format, colour = format, fill = format))
            + geom_histogram() + ylab("Count") + xlab("Format") +
              labs(title = plot.title))
      idx <- idx+1
    }
  }
}
```
