SI618 Group Project
========================================================

Chi-square Independence Test Analysis
--------------------------------------------------------

### Statistical Package Format and Accessor (Bots/Humans)

H<sub>0</sub>:  There **is no** association between statistical package format and accessor in the population of downloads in ICPSR web logs (they are independent).

H<sub>A</sub>:  There **is** an association between statistical package format and accessor in the population of downloads in ICPSR web logs.

```{r echo=FALSE}
library(DBI)
library(RSQLite)
library(stringr)
library(plyr)
library(ggplot2)

# Connect to the database...
driver<-dbDriver("SQLite")
connect<-dbConnect(driver, dbname = "../db/weblogs.db")

# Retrieve all weblog entries for stat package reports and define factors...
log.entries = dbGetQuery(connect, "
select wl.timestamp, wl.study, wl.browser_type, c.description || ',' as classification,
  case
    when wl.format like '%spss%' then 'SPSS'
    when wl.format like '%stata%' then 'STATA'
    when wl.format like '%sas%' then 'SAS'
    when wl.format = 'ascii' or wl.format = 'delimited' or wl.format = 'other' then 'DELIMITED'
    else wl.format
  end as format,
  case
    when (browser_type like 'Mozilla%' or browser_type like 'Opera%' or browser_type like 'Safari%' or browser_type like 'Nokia%' or browser_type like 'BlackBerry%') and
         (browser_type not like '%Yandex%' and browser_type not like '%spider%' and browser_type not like '%Babya Discover%'
         and browser_type not like '%bot%' and browser_type not like '%Daumoa%' and browser_type not like '%crawler%' and browser_type not like '%LinkCheck%'
         and browser_type not like '%WebCorp%' and browser_type not like '%Yahoo! Slurp%' and browser_type not like '%Genieo%') then 'Human'
         when browser_type like '%CopperEgg%' or browser_type like '%InterMapper%' then 'ICPSR-Bot'
    else 'Bot'
  end as accessor
from weblog_downloads wl inner join classifications c on wl.study = c.study_id
")

log.entries$format <- factor(log.entries$format)
log.entries$accessor <- factor(log.entries$accessor)
log.entries$classification <- factor(log.entries$classification)

# Perform a chi-square test on format vs accessor, report results...
ct <- chisq.test(log.entries$format, log.entries$accessor)
ct
```

Reject H<sub>0</sub>.  There is statistically significant evidence to support an association between statistical package format and accessor in the population of downloads in ICPSR web logs.  Residual value summary:

```{r echo=FALSE}
# Graph residuals...
res <- as.data.frame(ct$residuals)
the.plot <- ggplot(res, aes(log.entries.format, Freq, color = log.entries.accessor, group = log.entries.accessor)) +
  geom_line() + ylab("Residual") + xlab("Download Format") +
  guides(col=guide_legend(title = "Accessor"))
the.plot + labs(title = "Residuals of Format v Accessor")
the.plot + labs(title = "Residuals of Format v Accessor (Scale-adjusted)") + scale_y_continuous(limits = c(-100, 100))
```

### Statistical Package Format and Classification

H<sub>0</sub>:  There **is no** association between statistical package format and classification in the population of downloads in ICPSR web logs (they are independent).

H<sub>A</sub>:  There **is** an association between statistical package format and classification in the population of downloads in ICPSR web logs.

```{r echo=FALSE}
# Perform a chi-square test on format vs classification, report results...
log.entries$top_classification <- substr(log.entries$classification, start = 1, stop = str_locate(log.entries$classification, ",")-1)
log.entries$top_classification <- factor(log.entries$top_classification)
ct <- chisq.test(log.entries$format, log.entries$top_classification)
ct
```

Above warning due to some expected-count cells having value < 1.  Assumptions to conduct a Chi-square independence test are not met.  Here are residual values anyway:

```{r echo=FALSE, fig.width=15}
# Graph residuals...
res <- as.data.frame(ct$residuals)
the.plot <- ggplot(res, aes(log.entries.format, Freq, 
                      color = log.entries.top_classification,
                      group = log.entries.top_classification)) +
  geom_line() + ylab("Residual") + xlab("Download Format") +
  guides(col=guide_legend(title = "Classification"))
the.plot + labs(title = "Residuals of Format v Accessor")
the.plot + labs(title = "Residuals of Format v Accessor (Scale-adjusted)") + scale_y_continuous(limits = c(-100, 100))
```
